\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
\label{chap:abstract}

In this study, I introduce $Log_A$T, a comprehensive family of languages specifically developed for the purpose of reasoning about temporal ontologies,
with a particular focus on states and events.
This innovative framework goes beyond the limitations of traditional languages by adopting a purely term-based approach,
rendering $Log_A$T an algebraic language.

The primary objective of this thesis is to delve into the analysis and understanding of punctual and durative events within the context of $Log_A$T.
By thoroughly investigating these event types, I aim to provide valuable insights into their representation and integration within the framework.
This research endeavor serves as a significant extension to the existing $Log_A$S language, as it allows for the seamless incorporation and reasoning of events alongside states.

Furthermore, this study addresses the crucial task of establishing a comprehensive framework that accounts for the intricacies of events. By carefully examining the relationships between events and states, this research aims to contribute to the development of an enriched $Log_A$S language that can effectively capture and reason about the dynamic nature of events.

By expanding the scope and capabilities of $Log_A$S through the introduction of $Log_A$T, this research strives to provide a valuable contribution to the field of temporal ontologies.
Through the exploration and analysis of punctual and durative events, this study not only enhances our understanding of these concepts but also enables more sophisticated and comprehensive reasoning within the extended $Log_A$S framework.